---
title: "Standardized Residuals"
output: pdf_document
fontsize: 12pt
geometry: margin=1in
---
## Definition
**Standardized residuals** are the fitted model residuals divided by the residual standard deviation where the residual standard deviation is the square root of the Mean Square Error of the model. 

**Equations**
$$s_i = \frac{e_i}{\sqrt{MSE}} = \frac{e_i}{\hat{Var}(e_i)}$$
$$MSE = \frac{\sum(Y_i-\hat{Y_i})^2}{n-p}$$

## Example Plot
```{r echo=FALSE, fig.height=3.5, fig.width=6.5}
data(mtcars)
model <- lm(mpg ~ ., data=mtcars)
stdResiduals <- rstandard(model)
plot(stdResiduals, ylab="Standardized Residuals", main="Standardized Residuals for mtcars Model\n mpg ~ .")
abline(0,0)
abline(h=2, lty = 3)
abline(h=-2, lty = 3)
```

## Interpretation and Use
Standardized residuals allow for comparing on the standard scale ($\sigma^2=\sigma=1$) where values exceeding +-2 on the scale may indicate an unusual occurence for further investigation. 

**Problems**
Since standardized residuals derive from $e_i = y_i -\hat{y}_i$, a $y_i$ with high leverage will drag the regression towards it, influencing the residual estimate of $y_i$

**Solutions and Further Avenues**  
*Studentized Residuals -* fit the regression with $y_i$ excluded such that the residual becomes $y_i - \hat{y}_{i,(-i)}$ where $\hat{y}_{i,(-i)}$ denotes the regression line fit minus the point, $y_i$.

## R Code
```{r}
data(mtcars)
model <- lm(mpg ~ ., data=mtcars)
stdResiduals <- rstandard(model)
#plot(stdResiduals, ylab="Standardized Residuals", 
#        main="Standardized Residuals for mtcars Model\n mpg ~ .")
#abline(0,0)
#abline(h=2, lty = 3)
#abline(h=-2, lty = 3)
```
