---
title: "DFBETAS"
output: pdf_document
fontsize: 12pt
geometry: margin=1in
---

```{r echo=FALSE}
library(MASS)
library(car)
```

## Definition
**DFBETAS** is a post-regression tool for measuring the influence of outlying data points. It measures the effect on the regression coefficient for a variable of deleting each data observation. This difference is scaled using the standard error when the observation is omitted. When observations exceed the DFBETA cut-off value ($\frac{2}{\sqrt{n}}$), they should be investigated further as the observation could be overly influential on the estimated coefficent. 

**Equations**
      $$DFBETAS_{ij} = \frac{\hat{\gamma_i}-\hat{\gamma_{i(-j)}}}{se(\hat{\gamma_{i(-j)}})}$$

## Example DFBETAS Plots for Swiss Data Set
```{r echo=FALSE, message=FALSE, fig.height=5}
# use swiss dataset
data(swiss)
model <- lm(Fertility ~ ., data=swiss)
dfbetasPlots(model, main="Plot of Swiss DFBETAS") 
```

## Interpretation and Use
DFBETAS is useful, especially in concert with DFFITS and Cook's Distance, to identify and investigate high influence points in a dataset. High influence by itself is not a valid reason to eliminate an observation but indicates that a researcher should evaluate the point for validity in the data.    
      
      
**Further Avenues**  
*DFFITS/Cook's Distance* - In conjunction with DFBETAS, DFFITS and Cook's Distance should corroborate any high influence points in the data. 

## R Code
```{r}
# use swiss dataset
data(swiss)
model <- lm(Fertility ~ ., data=swiss)
#dfbetasPlots(model, main="Plot of Swiss DFBETAS") 
```