---
title: "DFFITS"
output: pdf_document
fontsize: 12pt
geometry: margin=1in
---

```{r echo=FALSE}
library(MASS)
library(car)
library(ggplot2)
```

## Definition
**DFFITS** is a post-regression tool for measuring the influence of outlying data points. It measures the effect on the regression coefficient for a variable of deleting each data observation and "studentized" by dividing by the standard deviation. The cut-off for DFFITS is $2\sqrt{\frac{p}{n}}$.

**Equations**
      $$DFFITS = \frac{\hat{y_i}-\hat{y_{i(i)}}}{s_{(i)}(\sqrt{h_{ii}})}$$

## Example DFFITS Plots for Swiss Data Set
```{r echo=FALSE, message=FALSE, fig.height=4}
# use swiss dataset
data(swiss)
model <- lm(Fertility ~ ., data=swiss)
dffitsData <- as.data.frame(dffits(model) )
names(dffitsData) <- c("observations")
cutoff <- 2*sqrt(6/47) #2*sqrt(p/n)
plot(dffitsData$observations, main="DFFITS Plot", xlab="observations", ylab = "DFFITS")
abline(h=cutoff,col=4,lty=2)
abline(h=-cutoff,col=4,lty=2)
labels=row.names(swiss)
text(c(6,37,42,46,47),dffitsData[c(6,37,42,46,47),],labels=labels[c(6,37,42,46,47)])
```

## Interpretation and Use
DFFITS is useful, especially in concert with DFBETAS and Cook's Distance, to identify and investigate high influence points in a dataset. High influence by itself is not a valid reason to eliminate an observation but indicates that a researcher should evaluate the point for validity in the data.    
      
      
**Further Avenues**  
*DFBETAS/Cook's Distance* - In conjunction with DFFITS, DFBETAS and Cook's Distance should corroborate any high influence points in the data. 

## R Code
```{r}
# use swiss dataset
# data(swiss)
# model <- lm(Fertility ~ ., data=swiss)
# dffitsData <- as.data.frame(dffits(model) )
# names(dffitsData) <- c("observations")
# cutoff <- 2*sqrt(6/47) #2*sqrt(p/n)
# plot(dffitsData$observations, main="DFFITS Plot", xlab="observations", ylab = "DFFITS")
# abline(h=cutoff,col=4,lty=2)
# abline(h=-cutoff,col=4,lty=2)
# labels=row.names(swiss)
# text(c(6,37,42,46,47),dffitsData[c(6,37,42,46,47),],labels=labels[c(6,37,42,46,47)])
```