---
title: "PRESS Statistic"
output: pdf_document
fontsize: 12pt
geometry: margin=1in
---
```{r echo=FALSE}
library(MASS)
```

## Definition
The **PRESS statistic** is the predicted sum of squares. This statistic provides an idea of the model's predictive ability. It is the sum of squares of the *prediction* error. For each $y_i$, the fitted value $\hat{y_i}$ is obtained from the remaining n-1 observations. Think of the n-1 observations as the training data for a single test data point for prediction. The sum of these squared errors, $y_i - \hat{y_i}$ is the PRESS statistic.  

**Equations**
      $$PRESS = \sum_{i=1}^{n}(y_i - \hat{y}_{i,-i})^2$$

## Example PRESS Stats
```{r echo=TRUE, message=FALSE}
library(MPV)
# use swiss dataset
data(swiss)
model <- lm(Fertility ~ ., data=swiss)
model2 <- update(model, Fertility ~ Agriculture + Education)
model3 <- update(model, Fertility ~ Agriculture + Education + Examination)
PRESSstat <- PRESS(model)
PRESSstat2 <- PRESS(model2)
PRESSstat3 <- PRESS(model3)
```

Model  | PRESS Statistic
:-------------:     | :-------------:|
Full Model        | 2815
Nested Model 1    | 4412  
Nested Model 2    | 3757

## Interpretation and Use
The PRESS statistic is useful, especially in conjunction with the studentized residuals (since they are based on the PRESS statistic instead of the sum of squares). Among similar or nested models, the PRESS statistic helps in evaluating the predictive abilities of the models. A lower PRESS is better among comparable models. Compared to standarized residuals, the PRESS statistic can help identify an overfitted model since the PRESS statistic will be larger since it evaluates the jackknifed residuals ("leave one [observation] out").   
      
      
**Further Avenues**  
*AIC/BIC* - The PRESS statistic works well in conjunction with the Akaike and Bayesian Information Criteria statistics (AIC/BIC). They measure the relative loss of information between nested models and are useful in determining if simpler models outperform more complex models. 

## R Code
```{r}
# alternative method to manually calculate the PRESS statistic 
# without the MPV library
pr <- residuals(model)/(1 - lm.influence(model)$hat)
PRESS <- sum(pr^2)
# identical result as the PRESS call to the full model above 
PRESS
```




