---
title: "Studentized Residuals"
output: pdf_document
fontsize: 12pt
geometry: margin=1in
---
```{r echo=FALSE}
library(MASS)
library(car)
```

## Definition
**Studentized residuals** are the standardized "jackknifed" residuals. "Jackknifing" is the process of *leave one out* such that the studentized residual is the residual where $y_i$ is compared to the fitted model point $\hat{y_i}$ where $y_i$ is not included (equation 1). In equation 1, the denominator is no longer the Mean Square Error, but the Predicted Sum of Squares, or PRESS statistics. NOTE: $h_{ii}$ is called "leverage" in that it measures the ability of a point to change the regression coefficient (slope) of the model. The greater the effect of moving $y_i$ on the regression slope, the greater its "leverage."

**Equations**
      $$t_i = \frac{\hat{e_i}}{\hat{\sigma}\sqrt{1-h_{ii}}}$$
      $$h_{ii} = \frac{d\hat{y_i}}{dy_i}$$
      
      
## Example Plots
```{r echo=FALSE, fig.height=3.5, fig.width=7}
data(swiss)
model <- lm(Fertility ~ ., data=swiss)
studentResids <- studres(model)
par(mfrow=c(1,2))
plot(studentResids, ylab="Studentized Residuals", main="Studentized Residuals for\n swiss Model Fertility ~ .")
abline(0,0)
abline(h=2, lty = 3)
abline(h=-2, lty = 3)
hist(studentResids, freq=FALSE, main="Distribution of\nStudentized Residuals", xlab="Residuals", ylim=c(0,0.5))
xfit<-seq(min(studentResids),max(studentResids),length=40) 
yfit<-dnorm(xfit) 
lines(xfit, yfit)
```

## Interpretation and Use
Studentized residuals allow for evaluating how much of an outlier or leverage point that each $y_i$ is in the data. It serves two purposes in diagnositics. First, the values of $t_i$ indicate if any points are particularly influential in the model if their values lie outside +-2. Second, the distribution of the studentized residuals can indicate if the data are non-normal when the distribution does not follow a normal distribution.  
      
**Further Avenues**  
*QQ Plots* - Studentized residuals provide a good method to identify if the residuals are non-normal. A QQ plot will better indicate if this is true for determining model goodness-of-fit. 

## R Code
```{r}
library(MASS)
library(car)
data(swiss)
model <- lm(Fertility ~ ., data=swiss)
studentResids <- studres(model)
#plot(studentResids, ylab="Studentized Residuals", 
#        main="Studentized Residuals for\n swiss Model Fertility ~ .")
#abline(0,0)
#abline(h=2, lty = 3)
#abline(h=-2, lty = 3)
#hist(studentResids, freq=FALSE, 
#        main="Distribution of\nStudentized Residuals", 
#        xlab="Residuals", ylim=c(0,0.5))
#xfit<-seq(min(studentResids),max(studentResids),length=40) 
#yfit<-dnorm(xfit) 
#lines(xfit, yfit)
```
